### 2021-10-27 Papers/WUT-Master-Thesis/tex/7-wnioski.tex

\section{Wnioski}

\subsection{Kategoryzacja języka}