### 2022-07-03 Papers/ICCC-2023/sections/realted-work.tex
.

\section{Related work}
\label{related-work}

Many currently used programming languages allow the programmer to execute some code at compile time, including C\# \cite{csharp:source_generators,roslyn}, Rust \cite{rust, klabnik2019rust} and C++ \cite{ISO:cpp20}.