### 2022-11-01 Papers/ICCC-2023/sections/realted-work.tex

\section{Related work}
\label{related-work}

Many currently used programming languages allow the programmer to execute some code at compile time, for example: C\# \cite{csharp:source_generators,roslyn}, Rust \cite{rust, klabnik2019rust} and C++ \cite{ISO:cpp20}.

Rust language compiler is the most similar in capability, to what was demonstrated with C-=-1, as a feature of CTFEF.
It allows the user to write code that performs some transformations of the program, and interact with the environment during the build process.
The two relevant features are build scripts and macros.
Rust macros, similar to the classic C-style preprocessor macros, generate additional code as text.
The major difference between the systems is that Rust macros operate on tokens, rather than text, and use syntax similar to regular Rust code.
This mechanism is powerful, allowing the user to rewrite code to avoid repetition, simplify certain tasks and create new syntax.
They are however unable to reflect on the program as a whole or obtain some information available to the compiler.

Build scripts, on the other hand, are programs that execute before the compilation of the main package.
They prepare the environment for building the program, for example, compile external dependencies.
The structure of the program being compiled, is not available for build scripts.
Build scripts accessing such data, would have to analyze the code by themselves, without any compiler assistance, and that makes automatic generation of bindings for other programming languages very difficult.