### 2023-01-06 Papers/ENASE-2023/c-=-1.tex
fixes

\documentclass[a4paper,twoside]{article}
\usepackage[OT1]{fontenc}

\usepackage{epsfig}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{apalike}
\usepackage[bottom]{footmisc}
\usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
\lstset{tabsize=2}
\begin{document}

\title{Compilation through interpretation}

\author{
	\authorname{
		Adam Grabski\sup{1}
		\orcidAuthor{0000-0002-2894-5976}, Ilona Bluemke\sup{1}\orcidAuthor{0000-0002-2894-5976}
	}
	\affiliation{\sup{1}Department of Electronics and Information Technology, Warsaw University of Technology, Nowowiejska 15/19, Warsaw, Poland}
	\email{adam.grabski@stud.pw.edu.pl, Ilona.Bluemke@pw.edu.pl}
}

\keywords{compiler, interpreter, metaprogramming}

\abstract{\input{sections/abstract.tex}}

\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill

\input{sections/introduction.tex}

\input{sections/realted-work.tex}

\input{sections/design.tex}

\input{sections/implementation.tex}

\input{sections/conclusions.tex}

%% Bibliography

\bibliographystyle{apalike}
{\small
	\bibliography{bibliography}}
\end{document}