### 2023-05-21 Papers/FEDCIS-2023/cm1.tex
added fedcis paper

% This file demonstrates how to use the IEEEConf LaTeX2e macro package,
% to prepare a manuscript for proceedings on CD of the conference
% FedCSIS
%
\documentclass[conference]{IEEEtran}

\IEEEoverridecommandlockouts

% This package serves to balance the column lengths on the last page of the document.
\usepackage{pbalance}

%% to enable \thank command
\IEEEoverridecommandlockouts
% to typeset algorithms
\usepackage{algorithmic}
\usepackage{algorithm}
% to make an accent \k be available
\usepackage[OT4,T1]{fontenc}
% provides various features to facilitate writing math formulas and to improve the typographical quality of their output.
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
% por urls typesetting and breaking
\usepackage{url}
% for vertical merging table cells
\usepackage{multirow}
\usepackage{flushend}

% define environments for remarks and examples
\newtheorem{remark}{Remark}[section]
\newtheorem{example}[remark]{Example}

\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{graphicx}
\usepackage{listings}
\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
\lstset{framextopmargin=50pt,frame=bottomline}
\bibliographystyle{acm}

\begin{document}

\title{Compilation through interpretation}

\author{
	\IEEEauthorblockN{Adam Grabski}
	\IEEEauthorblockA{0000-0002-6283-8461\\
		Warsaw Univeristy of Technology\\
		in Warsaw\\
		ul.\ Nowowiejska 15/19, 00-665 Warsaw, Poland\\
		Email: adam.grabski@stud.pw.edu.pl}
	\and
	\IEEEauthorblockN{Ilona Bluemke}
	\IEEEauthorblockA{0000-0002-2894-5976\\
		Warsaw Univeristy of Technology\\
		in Warsaw\\
		ul.\ Nowowiejska 15/19, 00-665 Warsaw, Poland\\
		Email: Ilona.Bluemke@pw.edu.pl}
}

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\input{sections/introduction.tex}

\input{sections/realted-work.tex}

\input{sections/design.tex}

\input{sections/implementation.tex}

\input{sections/conclusions.tex}

%% Bibliography
\bibliography{bibliography}

\end{document}