### 2023-07-17 Papers/FEDCIS-2023/cm1.tex
corrections from FEDCIS review process

% This file demonstrates how to use the IEEEConf LaTeX2e macro package,
% to prepare a manuscript for proceedings on CD of the conference
% FedCSIS
%
\documentclass[conference]{IEEEtran}

\IEEEoverridecommandlockouts

% This package serves to balance the column lengths on the last page of the document.
\usepackage{pbalance}

%% to enable \thank command
\IEEEoverridecommandlockouts
% to typeset algorithms
\usepackage{algorithmic}
\usepackage{algorithm}
% to make an accent \k be available
\usepackage[OT4,T1]{fontenc}
% provides various features to facilitate writing math formulas and to improve the typographical quality of their output.
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
% por urls typesetting and breaking
\usepackage{url}
% for vertical merging table cells
\usepackage{multirow}
\usepackage{flushend}

% define environments for remarks and examples
\newtheorem{remark}{Remark}[section]
\newtheorem{example}[remark]{Example}

\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{graphicx}
\usepackage{listings}
\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
\lstset{framextopmargin=50pt,frame=bottomline}
\bibliographystyle{acm}

\begin{document}

\title{Compilation through interpretation}

\author{
	\IEEEauthorblockN{Adam Grabski}
	\IEEEauthorblockA{0000-0002-6283-8461\\
		Warsaw Univeristy of Technology\\
		in Warsaw\\
		ul.\ Nowowiejska 15/19, 00-665 Warsaw, Poland\\
		Email: adam.gr@outlook.com}
	\and
	\IEEEauthorblockN{Ilona Bluemke}
	\IEEEauthorblockA{0000-0002-2894-5976\\
		Warsaw Univeristy of Technology\\
		in Warsaw\\
		ul.\ Nowowiejska 15/19, 00-665 Warsaw, Poland\\
		Email: Ilona.Bluemke@pw.edu.pl}
}

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\begin{abstract}

	As static metaprogramming is becoming more relevant, compilers must adapt to accommodate them. This requires exposing more information about the code, from the compiler to the programmer as well as more powerful compile-time function execution capabilities. The Interpreter component of a compiler therefore becomes more important.

	In this paper a novel approach to compiler architecture that places the Interpreter as the central component of the compiler is proposed.
	Translation of user code into the executable form is done by an Interpreter, written in the target language. The data structures of Interpreter are accessible also to the programmer. Such solution significantly improves flexibility and extensibility of the compiler and enables execution of any code at the compile-time.

	Compile Time Function Execution (CTFE) First pattern was designed for low-level, non-garbage-collected, reflection-enabled language C-=-1. This is a new programming language, which allows the programmer to execute any code at the compile time, as well as to analyze and modify the program structure. Grace to the extensibility of the designed compiler, programs written in C-=-1 can also generate marshalling bindings for other languages and support a variety of programming paradigms.

	The significant flexibility and extensibility of CTFEF caused severe problems in compiler construction. The compiler appeared very complex, its parts, especially Interpreter, were very difficult to debug. Compiler operates on inflexible data structures, accessible also for a programmer. They are therefore a part of the compiled languages standard library and backwards compatibility must be maintained.

\end{abstract}

\input{sections/introduction.tex}

\input{sections/realted-work.tex}

\input{sections/design.tex}

\input{sections/implementation.tex}

\input{sections/conclusions.tex}

%% Bibliography
\bibliography{bibliography}

\end{document}