### 2023-08-19 Papers/FEDCIS-2023-presentation/sections/motivation.tex

\section{Motivation}

\begin{frame}
	\frametitle{Metaprogramming}

	\begin{itemize}
		\item Metaprogramming - writing programs that analyse or modify themselves\begin{itemize}
			      \item Dynamic - metaprogramming done at run time
			      \item Static - metaprogramming done at compile time
		      \end{itemize}
		\item Makes code more expressive and resilient to change
		\item Boosts productivity
		\item Can improve performance
		\item Mostly present in garbage-collected languages at runtime\begin{itemize}
			      \item C\#
			      \item Java
			      \item Scripting languages such as Javscript or python
		      \end{itemize}
	\end{itemize}

\end{frame}

\begin{frame}
	\frametitle{Static metaprogramming}

	\begin{itemize}
		\item Until recently a niche tool \begin{itemize}
			      \item C style preprocessor macros
			      \item D templates, compile time function execution and string mixins
		      \end{itemize}
		\item Significant developments since 2011\begin{itemize}
			      \item C++ SFINAE template metaprogramming (standardized 2011)
			      \item Rust procedural macros (introduced with the language in 2015)
			      \item C\# source generators (introduced in 2020)
		      \end{itemize}

		\item Significant current interest and more development in the future\begin{itemize}
			      \item Migration of many typical dynamic metaprogramming activities to compile time in C\#
			      \item Introduction of compile time interceptors in C\#12
			      \item Continuous development of generic programming in C++
		      \end{itemize}
	\end{itemize}

\end{frame}