### 2023-08-28 Papers/FEDCIS-2023-presentation/cm1.tex
some stuff

\documentclass[10pt]{beamer} %Beamer

\usepackage{palatino} %font type
\usepackage{caption}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{csquotes}

\usepackage{appendixnumberbeamer} %enumerate each slide without counting the appendix

%These next packages are the useful for Physics in general, you can add the extras here.
\usepackage{amsmath,amssymb}
\usepackage{slashed}
\usepackage{dirtree}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{pgfplots}
\usepgfplotslibrary{dateplot}
\usepackage{geometry}
\usepackage{xspace}
\usepackage{color}

\lstloadlanguages{C,C++,csh,Java}
\usefonttheme{metropolis} %Type of slides
\usefonttheme[onlymath]{serif} %font type Mathematical expressions
\usetheme[progressbar=frametitle,titleformat frame=smallcaps,numbering=counter]{metropolis} %This adds a bar at the beginning of each section.
\useoutertheme[subsection=false]{miniframes} %Circles in the top of each frame, showing the slide of each section you are at

\definecolor{red}{rgb}{0.6,0,0}
\definecolor{blue}{rgb}{0,0,0.6}
\definecolor{green}{rgb}{0,0.8,0}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}

\lstset{
language=csh,
basicstyle=\footnotesize\ttfamily,
numbers=left,
numberstyle=\tiny,
numbersep=5pt,
tabsize=2,
extendedchars=true,
breaklines=true,
frame=b,
stringstyle=\color{blue}\ttfamily,
showspaces=false,
showtabs=false,
xleftmargin=17pt,
framexleftmargin=17pt,
framexrightmargin=5pt,
framexbottommargin=4pt,
commentstyle=\color{green},
morecomment=[l]{//}, %use comment-line-style!
morecomment=[s]{/*}{*/}, %for multiline comments
showstringspaces=false,
morekeywords={ abstract, event, new, struct,
as, explicit, null, switch,
base, extern, object, this,
bool, false, operator, throw,
break, finally, out, true,
byte, fixed, override, try,
case, float, params, typeof,
catch, for, private, uint,
char, foreach, protected, ulong,
checked, goto, public, unchecked,
class, if, readonly, unsafe,
const, implicit, ref, ushort,
continue, in, return, using,
decimal, int, sbyte, virtual,
default, interface, sealed, volatile,
delegate, internal, short, void,
do, is, sizeof, while,
double, lock, stackalloc,
else, long, static,
enum, namespace, string},
keywordstyle=\color{cyan},
identifierstyle=\color{red},
backgroundcolor=\color{white},
}
\usepackage{biblatex}
\usepackage{caption}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{blue}{\parbox{\textwidth}{\hspace{15pt}#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize}}
\newcommand{\themename}{\textbf{\textsc{bluetemp}\xspace}}%metropolis}}\xspace}

\title{CTFEF }
\subtitle{Building a compiler for metaprogramming languages}
\author[name]{
	mgr. inż. Adam Grabski\\
	dr. hab. inż Ilona Bluemke
}

\institute[uni]{ Department of Electronics and Information Technology \\
Warsaw Univeristy of Technology
}
\date{\today} %Here you can change the date
\begin{document}
{

\maketitle
}%This is the colour of the first slide. bg= background and fg=foreground

\begin{frame}{Agenda}
	\setbeamertemplate{section in toc}[sections numbered] %This is numbering the sections
	\tableofcontents[subsectionstyle=hide/hide/hide] %You can comment this line if you want to show the subsections in the table of contents
\end{frame}

\input{sections/motivation.tex}
\input{sections/CTFEF.tex}

\end{document}