### 2023-08-28 Papers/FEDCIS-2023-presentation/sections/CTFEF.tex

\section{CTFEF}

\begin{frame}
    \frametitle{CTFEF compiler architecture}
    \begin{figure}

        \begin{center}
            \includegraphics[height=6cm]{pictures/compiler-structure.png}
        \end{center}
    \end{figure}

\end{frame}

\begin{frame}
    \frametitle{Frontend}

    \begin{itemize}
        \item Translates source code into Intermidiate Representation\begin{itemize}
            \item Parse
            \item Validate syntax
            \item Inline operators with special syntax (ex: \lstinline|typeof|)
        \end{itemize}
        \item 
    \end{itemize}

\end{frame}