### 2023-08-30 Papers/FEDCIS-2023-presentation/sections/C-=-1.tex
more presentation stuff

\section{C-=-1}

\begin{frame}
    \frametitle{Type System}

\end{frame}

\begin{frame}
    \frametitle{Semantic model}

\end{frame}

\begin{frame}
    \frametitle{Compiler structure}

\end{frame}