### 2023-09-04 Papers/FEDCIS-2023-presentation/sections/Conclusions.tex
more presentation stuff

\section{Conclusions}

\begin{frame}
    \frametitle{Conclusions}

\end{frame}

\begin{frame}
    \frametitle{Conclusions}

    \begin{itemize}
        \item C-=-1 performance is unsatisfactory\begin{itemize}
            \item ~10 minutes to compile 200 lines of code
        \end{itemize}
        \item Performance of the compiler was not a subject of the research, a lot of improvements are possible\begin{itemize}
            \item Currently only debug build works
            \item Interpreter data structures are inefficient (significant amount of indirection when accessing object members)
        \end{itemize}
        \item Further work is needed
    \end{itemize}

\end{frame}