### 2023-09-06 Papers/FEDCIS-2023-presentation/sections/Conclusions.tex
final version of fedcis presentation?

\section{Conclusions}

\begin{frame}
    \frametitle{Conclusions}

    \begin{itemize}
        \item CTFEF architecture is flexible \begin{itemize}
            \item Several language features from other languages implemented as library features in C-=-1
            \item Generating source code of other languages at compiletime
        \end{itemize}
        \item Parts of compiler written in target language on very early stage
        \item Large amount ow work devoted to parts of compiler related to interpreting code\begin{itemize}
            \item Interpreter
            \item Defining primitive operations
            \item Bindings to LLVM
        \end{itemize}
    \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{Conclusions}

    \begin{itemize}
        \item C-=-1 compiler performance is unsatisfactory\begin{itemize}
            \item ~10 minutes to compile 200 lines of code
        \end{itemize}
        \item Performance of the compiler was not a subject of the research, a lot of improvements are possible\begin{itemize}
            \item Currently only debug build works
            \item Interpreter data structures are inefficient (significant amount of indirection when accessing object members)
        \end{itemize}
        \item Further work is needed
    \end{itemize}

\end{frame}